---
title: "R Markdown, Knitr Assignment"
author: Pamela
date: Dec 2017
output: md_document:  variant: markdown_github
---
#R Markdown, Knitr Assignment

##### 1.Loading and preprocessing the data
```{r readcsv, echo= TRUE}
MyData <- read.csv(file="activity.csv", header=TRUE, sep=",")
MyData_New <- subset(MyData, is.na(MyData$steps) == FALSE)
```

##### 2.Mean total number of steps taken per day
```{r meantotal}
require(stats)
MyData_Daysum <-tapply(MyData_New$steps, MyData_New$date, sum)
hist(MyData_Daysum, col="lightgreen", xlab ="Sum of Steps", main = "Histogram of Steps per Day")
MyData_Daymean <-tapply(MyData_New$steps, MyData_New$date, mean, na.rm = TRUE)
MyData_Daymed <-tapply(MyData_New$steps, MyData_New$date, median , na.rm = TRUE)
```
######The mean steps per day: `r MyData_Daymean`
######The median steps per day: `r MyData_Daymed`

#####3. Average daily activity pattern
```{r AvgActPat}
MyData_Intrvmean <-tapply(MyData_New$steps, MyData_New$interval, mean)
plot(MyData_Intrvmean, type = "l",xlab ="Interval",ylab ="Mean Steps" ,xaxt = "n", col="blue" )
axis(1, at=1:length(names(MyData_Intrvmean)), labels =names(MyData_Intrvmean) )

#Maximum average interval mean
MyData_Intrvmean_DF<- data.frame(interval=names(MyData_Intrvmean), value=MyData_Intrvmean)
x <-MyData_Intrvmean_DF[which.max(MyData_Intrvmean_DF$value),]
```
######The maximum  of average steps per day  `r x$value` falls at  `r x$interval`th  interval

#####4. Imputing missing values

######Count of missing values `r nrow(MyData[!complete.cases(MyData),])` using nrow(MyData[!complete.cases(MyData),])

```{r}
MyData2 <- MyData
cleanBase2 <- MyData2[!is.na(MyData2$steps),]
missingData <- is.na(MyData2$steps)
meanVals <- tapply(cleanBase2$steps, cleanBase2$interval, mean, na.rm=TRUE, simplify=TRUE)
MyData2$steps[missingData] <- meanVals[as.character(MyData2$interval[missingData])]
#count of missing values in new data
nrow(MyData2[!complete.cases(MyData2),])

#imputed histogram
MyData2_Daysum <-tapply(MyData2$steps, MyData2$date, sum)
hist(MyData2_Daysum, col="skyblue", xlab ="Sum of Steps", main = "Histogram of Steps per Day-Missing values imputed")
MyData2_Daymean <-tapply(MyData2$steps, MyData2$date, mean)
MyData2_Daymed <-tapply(MyData2$steps, MyData2$date, median)

#Calculate and report the mean and median total number of steps taken per day. 
MyData_Daysum_DF<- data.frame(day=names(MyData_Daysum), value=MyData_Daysum)
MyData2_Daysum_DF<- data.frame(day=names(MyData2_Daysum), value=MyData2_Daysum)

D <-median(MyData_Daysum_DF$value, na.rm =  TRUE)
D2 <-median(MyData2_Daysum_DF$value, na.rm =  TRUE)
S <- round((sum(MyData_Daysum_DF$value, na.rm =  TRUE))/61,2)
S2 <-round((sum(MyData2_Daysum_DF$value, na.rm =  TRUE))/61,2)

```

###### Mean before imputing is `r S` and  after imputing is `r S2`.  While median before was `r D` and after is `r D2`. Since median is equal to mean after inputing we can say this data has stronger central tendency than the one before imputing 

#####Are there differences in activity patterns between weekdays and weekends?
```{r}
MyData2_Wk <- MyData2
MyData2_Wk$WkDay_WkEnd <-ifelse( weekdays(as.Date(MyData2_Wk$date)) == "Saturday" | weekdays(as.Date(MyData2_Wk$date)) == "Sunday", "WkEnd", "WkDay")
MyData2_WkDay <- MyData2_Wk[MyData2_Wk$WkDay_WkEnd == "WkDay", ]
MyData2_WkEnd <- MyData2_Wk[MyData2_Wk$WkDay_WkEnd == "WkEnd", ]
MyData2_WkDay <- tapply(MyData2_WkDay$steps, MyData2_WkDay$interval, mean, na.rm = TRUE)
MyData2_WkEnd <- tapply(MyData2_WkEnd$steps, MyData2_WkEnd$interval, mean, na.rm = TRUE)
##Plots of Weekday, Weekend
plot(MyData2_WkDay, type = "l",main ="Week Day",xlab="",ylab ="Daily Mean Steps" ,xaxt = "n", col="brown" )
axis(1, at=1:length(names(MyData2_WkDay)), labels =names(MyData2_WkDay) )
plot(MyData2_WkEnd, type = "l",main ="Week End",xlab="",ylab ="Daily Mean Steps" ,xaxt = "n", col="darkgreen" )
axis(1, at=1:length(names(MyData2_WkEnd)), labels =names(MyData2_WkEnd) )
```

